---
title: "Amaizing  Baseline Survey_Final-fertilizer"
author: "Njambi"
format: html
editor: visual
---

## 

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Clear the environment

```{r}

rm(list = ls())

```

## Loading necessary libraries

```{r, include=FALSE}

library(tidyverse) # includes the dplyr, ggplot2, tidyr, readr, stringr, forcats, tibble packages
library(readxl)
library(Hmisc)
library(labelled)
library(magrittr)
library(datawizard)
library(gtsummary)
library(gt)
library(kableExtra)
library(flextable)
library(splitstackshape)
library(lubridate)
library(foreign)
library(haven)

#library(ggplot2)
#library(tidyr)
#library(tibble)
#library(dplyr)
#library(stringr)

```

## Set the working directory

```{r}
# Set working directory (Not necessary since the current wd is the correct wd)

setwd("C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2024\\AMAZING\\AmazingBaselineDataCleaning\\Amaizing  Baseline Survey_Final-fertilizer")
      
```

## Loading data

```{r}
# Read a csv file

amaizing_fertilizer_path <- "C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2024\\AMAZING\\aMaizingBaseline_Files\\Amaizing  Baseline Survey_Final-fertilizer.csv"

amaizing_fertilizer <- read_csv(amaizing_fertilizer_path)

dim(amaizing_fertilizer) # 

```

```{r}

names(amaizing_fertilizer)
```

```{r}
unique(amaizing_fertilizer$fertindex2)


```

```{r}

fertilizer_levels <- c(
  "CAN", "DAP", "NPK", "Rock Phosphate", "Urea", "Branded fertilizer blends (Yara)", "Lime", "Foliar Fertilizer", "None of the above", "Other Specify"
                         )  

amaizing_fertilizer <- amaizing_fertilizer %>%
  mutate(
    fertindex2 = factor(
      fertindex2, 
      levels = c(1:3, 5:9, -666, -222), 
      labels = fertilizer_levels
    )
  
  ) 

```

```{r}
unique(amaizing_fertilizer$fertindex2)


```

```{r}

unique(amaizing_fertilizer$fertname)
```

The column `fertname` included text entry from "other specify", which has not been cleaned and harmonized.

```{r}


amaizing_fertilizer <- amaizing_fertilizer %>%
  mutate(
    fertname = factor(fertname)
  )
```

```{r}
unique(amaizing_fertilizer$fertname)

```

```{r}
unique(amaizing_fertilizer$f8)

```

```{r}
unique(amaizing_fertilizer$f9)

```

```{r}

unitfert_levels <- c(
"50kg Bag", "25kg Bag", "10kg Bag", "Kg", "Liter", "Gorogoro (2kgs)", "Debe (16Kgs)", "Wheelbarrow", "Cart", "Grams", "Other, Specify" 
                         )  

amaizing_fertilizer <- amaizing_fertilizer %>%
  mutate(
    f9 = factor(
      f9, 
      levels = c(1:9, 12, -222), 
      labels = unitfert_levels
    )
  
  ) 

```

```{r}
unique(amaizing_fertilizer$f9)

```

```{r}
names(amaizing_fertilizer)
```

```{r}
unique(amaizing_fertilizer$f9_sp)

# Drop this column; it is empty
```

```{r}

amaizing_fertilizer <- amaizing_fertilizer %>%
  select(-f9_sp)

names(amaizing_fertilizer)
```

```{r}

column_labels <- c(

fertindex1 = "Fertilizer repeat iteration/index",
fertindex2 = "Fertilizer level/category selected",
fertname = "Fertilizer label/name corresponding to fertindex2",
f8 = "Quantity of fertname used in plotid",
f9 = "Fertilizer Unit  for fertname used  in plotid",
#f9_sp = "Please Specify",
PARENT_KEY = "Parent key",
KEY = "Key",
`SET-OF-fertilizer` = "Fertilizer repeat group ID"
)

```

```{r}

# Rename columns with characters not supported by stata

amaizing_fertilizer <- amaizing_fertilizer %>%
  
  rename(SET_OF_fertilizer = `SET-OF-fertilizer`
    
         )

```

```{r}
# Set variable/column labels
 
amaizing_fertilizer <- amaizing_fertilizer %>%
  
  set_variable_labels(.labels = column_labels)   # Requires package labelled
```

Using the function View() to open the new data in a spreadsheet reveals the column names and labels.

```{r}

# View(amaizing_fertilizer)

```

## Generate a data dictionary

```{r}
# Data dictionary

amaizing_fertilizer_dict <- look_for(amaizing_fertilizer, labels = TRUE, values = TRUE) %>%
  dplyr::select(-c(missing, value_labels))

  gt(amaizing_fertilizer_dict)

```

## Export data dictionary to CSV

```{r}

readr::write_csv(amaizing_fertilizer_dict, "amaizing_fertilizer_dict.csv")

```

## Export data to CSV

```{r}

readr::write_csv(amaizing_fertilizer, "amaizing_fertilizer.csv")

```

## Save as R data

```{r}

save(amaizing_fertilizer, file = "amaizing_fertilizer.RData")

#View(amaizing_fertilizer)
```

## Export data to SPSS

```{r}

write.foreign(amaizing_fertilizer, "amaizing_fertilizer.sav", "amaizing_fertilizer", package = "SPSS")

```

## Export data to STATA

```{r}

write_dta(amaizing_fertilizer, "amaizing_fertilizer.dta")

```
